Combinatorics of the geometry of Wilson loop diagrams

### 2019-01-10 General WLD combinatorics.tex
section 3 lemma refs, and the beginnings of a bibliography

Let $\binom{[n]}{k}$ be the set of all $k$-subsets of the cyclically ordered set $[n]$.  For each $j \in [n]$, we can define a total order $\leq_j$ on the interval $[n]$ by
\bas j <_j j+1 <_j \dots <_j n <_j 1 \dots <_j j-1\;.\eas
This in turn induces a total order on $\binom{[n]}{k}$, namely the lexicographic order with respect to $<_j$.  It also induces a separate partial order $\gale{j}$ on $\binom{[n]}{k}$ (the \textit{Gale order} [ref]), which is defined as follows: for 
\[A = \{a_1 <_j a_2 <_j \dots <_j a_k\} \text{ and } B = \{b_1 <_j b_2 <_j \dots <_j b_k\} \in \binom{[n]}{k},\] we say that
\[A \gale{j} B \text{ if and only if } a_r \leq_j b_r \text{ for all }1 \leq r \leq k.\]
For example, in $\binom{[6]}{3}$ we have $\{2,5,6\}\gale{2} \{2,6,1\}$ but $\{2,5,6\}\not\gale{2}\{3,4,6\}$.

@@

By [ref][Oh or Postnikov?], the Grassmann necklaces of type $(k,n)$ are in 1-1 correspondence with the positroid cells in $\Gr(k,n)$.  Further, if $(I_1, \dots, I_n)$ is the Grassmann necklace associated to a positroid $M = ([n],\cB)$, then by [ref][Oh] the bases of $M$ can be computed using the Gale order $\gale{i}$ for each $i \in [n]$:
\[\cB = \left\{J \in \binom{[n]}{k}\ :\ I_i \gale{i} J \ \forall i \in [n]\right\}.\]
We now describe an algorithm that constructs the Grassmann necklace of an admissible Wilson loop diagram from the diagram. The fact that Algorithm~\ref{alg:put GN on WLD} does construct the required Grassmann necklace is proved in Theorem~\ref{res:alg gives GN}.

@@

Given a propagator $p$ of an admissible WLD $W$ and a vertex $i$ in the support of $p$, define
\[J_p^{(W)}(i) = \{m \in [n] \ : \ I^{(W)}_m(p) = i \},\]
i.e. the set of indices $m$ for which Algorithm [ref] assigns the value $i$ to the propagator $p$ in $W$. The following lemma establishes that these sets behave in a simple and predictable manner, a fact which we will repeatedly use in subsequent proofs.

@@

The dimension of a positroid cell is the same as the number of plusses in the associated Le diagram ***cite something***, and so with Algorithm~\ref{alg:put GN on WLD} along with the algorithm for passing from a Grassmann necklace to a Le diagram (reversing Oh's algorithm) described by some of us in \cite{reversingOh}, we can, at the cost of a bit of messy grunt work, give a recursive proof of the $3|\cP|$-dimensionality that explicitly describes the changes in the positions of the plusses in the Le diagram.

@@

\begin{lem}\label{lem dihedral}
  If two Wilson loop diagrams differ by a dihedral transformation then their Le diagrams have same number of plusses.
\end{lem}

\note{Can I just cite something for this, or can one of you provide a quick geometric proof.  It would be a horrible mess to do by hand combinatorially}

@@

>>>>

Let $\binom{[n]}{k}$ be the set of all $k$-subsets of the cyclically ordered set $[n]$.  For each $j \in [n]$, we can define a total order $\leq_j$ on the interval $[n]$ by
\bas j <_j j+1 <_j \dots <_j n <_j 1 \dots <_j j-1\;.\eas
This in turn induces a total order on $\binom{[n]}{k}$, namely the lexicographic order with respect to $<_j$.  It also induces a separate partial order $\gale{j}$ on $\binom{[n]}{k}$ (the \textit{Gale order} \cite{Gale}), which is defined as follows: for 
\[A = \{a_1 <_j a_2 <_j \dots <_j a_k\} \text{ and } B = \{b_1 <_j b_2 <_j \dots <_j b_k\} \in \binom{[n]}{k},\] we say that
\[A \gale{j} B \text{ if and only if } a_r \leq_j b_r \text{ for all }1 \leq r \leq k.\]
For example, in $\binom{[6]}{3}$ we have $\{2,5,6\}\gale{2} \{2,6,1\}$ but $\{2,5,6\}\not\gale{2}\{3,4,6\}$.

@@

By \cite[Theorem 17.1]{Postnikov}, the Grassmann necklaces of type $(k,n)$ are in 1-1 correspondence with the positroid cells in $\Gr(k,n)$.  Further, if $(I_1, \dots, I_n)$ is the Grassmann necklace associated to a positroid $M = ([n],\cB)$, then by \cite[Theorem 8]{Oh} the bases of $M$ can be computed using the Gale order $\gale{i}$ for each $i \in [n]$:
\[\cB = \left\{J \in \binom{[n]}{k}\ :\ I_i \gale{i} J \ \forall i \in [n]\right\}.\]
We now describe an algorithm that constructs the Grassmann necklace of an admissible Wilson loop diagram from the diagram. The fact that Algorithm~\ref{alg:put GN on WLD} does construct the required Grassmann necklace is proved in Theorem~\ref{res:alg gives GN}.

@@

Given a propagator $p$ of an admissible WLD $W$ and a vertex $i$ in the support of $p$, define
\[J_p^{(W)}(i) = \{m \in [n] \ : \ I^{(W)}_m(p) = i \},\]
i.e. the set of indices $m$ for which Algorithm~\ref{alg:put GN on WLD} assigns the value $i$ to the propagator $p$ in $W$. The following lemma establishes that these sets behave in a simple and predictable manner, a fact which we will repeatedly use in subsequent proofs.

@@

The dimension of a positroid cell is the same as the number of plusses in the associated Le diagram \cite[Theorem 6.5]{Postnikov}, and so with Algorithm~\ref{alg:put GN on WLD} along with the algorithm for passing from a Grassmann necklace to a Le diagram (reversing Oh's algorithm) described by some of us in \cite{reversingOh}, we can, at the cost of a bit of messy grunt work, give a recursive proof of the $3|\cP|$-dimensionality that explicitly describes the changes in the positions of the plusses in the Le diagram.

@@

\begin{lem}\label{lem dihedral}
If two Wilson loop diagrams differ by a dihedral transformation then their Le diagrams have same number of plusses.
\end{lem}
\begin{proof}
By \cite[Proposition 17.10]{Postnikov}, the dimension of a positroid (and hence the number of plusses in its Le diagram) is $k(n-k) - A(\pi_W)$, where $A(\pi_W)$ denotes the number of alignments (see \cite[Figure 17.1]{Postnikov} and preceding discussion) of the decorated permutation $\pi_W$ of the positroid associated to $W$.

Since it is clear that dihedral transformations of a Wilson loop diagram $W$ correspond to dihedral transformations of the chord diagram representation of $\pi_W$, and that the number of alignments in a chord diagram is preserved under dihedral transformations, the result follows.
\end{proof}

@@

\begin{thebibliography}{10}

\bibitem{wilsonloop}
S. Agarwala, E. M. Amat.
\newblock Wilson loop diagrams and positroids.
\newblock {\em Commun. Math. Phys.} 2016.

\bibitem{reversingOh}
S. Agarwala, S. Fryer.
\newblock An algorithm to construct the Le diagram associated to a Grassmann necklace
\newblock \url{arXiv:1803.01726}

\bibitem{Gale}
D. Gale.
\newblock Optimal assignments in an ordered set: {A}n application of matroid theory.
\newblock {\em J. Combinatorial Theory}, 4: 176 - 180 (1968).

\bibitem{Oh}
S. Oh.
\newblock Positroids and {S}chubert matroids.
\newblock {\em J. Combin. Theory Ser. A}, 118(8):2426--2435, (2011)

\bibitem{Postnikov}
A.~Postnikov.
\newblock Total positivity, {G}rassmannians, and networks.
\newblock \url{arXiv:math/0609764}.

\end{thebibliography}
